---
title: "Sims Model"
author: "Zoilo Franco"
date: "2026-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
require(tidyverse)
```

# Introduction

Form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 90 for GRIP and 90 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

```{r}
simData <- read.csv("sims.csv")
newdata = data.frame(ARM=90, GRIP=90)
```

# Methods

Each model predicts SIMS (numerical) based on ARM or GRIP strength, both of these are numerical also.
    
WE use the numerical~numerical helper file
    
# Model 1 SIMS depends on ARM
```{r}
m1 = lm(SIMS~ARM,data=simData)
```

## Numerical Summary model 1
```{r}
summary.lm(m1)
```
    
Model 1 is y -4.09 + 0.054*ARM    
Adjusted R-squared is .467    
Both coefficients are significantly different from 0
    
## graphical desciption of SIMS~ARM

```{r}
plot(SIMS~ARM,data=simData)
abline(m1)
```
    
The scatterplot shows a positive linear relation between SIMS and ARM
    
## Prediction for ARM=90
```{r}
predict(m1, newdata, interval='prediction')
```
    

    
# Model 2 SIMS depends on GRIP
```{r}
m2 = lm(SIMS~GRIP,data=simData)
```

## Numerical Summary model 2
```{r}
summary.lm(m2)
```
    
Model 2 is y -4.80 + 0.045*GRIP    
Adjusted R-squared is .405    
Both coefficients are significantly different from 0

## graphical desciption of SIMS~GRIP

```{r}
plot(SIMS~GRIP,data=simData)
abline(m2)
```
    
The scatterplot shows a positive linear relation between SIMS and GRIP
    
## Prediction for GRIP=90
```{r}
predict(m2, newdata, interval='prediction')
```
    

    
# Model 3 SIMS depends on ARM+GRIP
```{r}
m3 = lm(SIMS~ARM+GRIP,data=simData)
```

## Numerical Summary model 3
```{r}
summary.lm(m3)
```
    
Model 3 is y -5.43 + 0.04 * ARM + 0.02 * GRIP
Adjusted R-squared is .535    
Both coefficients are significantly different from 0

## graphical desciption of SIMS~ARM+GRIP

```{r}
plot(SIMS~ARM+GRIP,data=simData)
abline(m3)
```
    
The scatterplot shows a positive linear relation between SIMS and ARM+GRIP
    
## Prediction for ARM=90 and GRIP=90
```{r}
predict(m3, newdata, interval='prediction')
```
    
# Comparison using Anova between Model 1 and Model 3
    
```{r}
anova(m1,m3)
```
    
The Anova shows that model 3 is much better than model 1. P-value is equal to 0.000004994 which is smaller than .05 so we have evidence their is a difference in the models. Because model 3 has a smaller RSS, it is the better model.
    
# Discussion and Conclusion
    
Using three possible linear models for SIMS, SIMS~ARM(model 1), SIMS~GRIP(model 2), and SIMS~ARM+GRIP(model 3).
    
The anova shoes that model 3 was better than 1.

    